Gaia's constraints on the extremely massive ($\sim$$150\,M_\odot$) runaway star candidate VFTS682

### 2018-06-11 VFTS682_APJL.tex
fixed typo in correlation coefficient in tab.2
reverted to non-normalized plots in fig.1, it is so much clearer for me what is happening

       \hline
           \multicolumn{5}{l}{\emph{Gaia's proper motion for VFTS 682
      and average of the region}} \\
      \hline
          &$\mu_\mathrm{RA}$&[$\masyr$] & $1.843\pm 0.070$
                     & (1) \\        
               &$\mu_\mathrm{DEC}$&[$\masyr$] & $0.786\pm 0.080$
                     &  (1) \\        
                 & $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$ &[$\masyr$] & $0.023$
                        & (1)  \\         
     %  \multicolumn{5}{l}{\emph{Gaia's average proper motion in the region}} \\         
       &$\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$&[$\masyr$] & $1.74\pm0.01$
                        & (2) \\
      &$\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$&[$\masyr$]
                & $0.70\pm0.02$ &  (2)\\
\hline

@@

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}
  \includegraphics[width=0.35\textwidth]{figures/angle_dist}
  \caption{Left panel: distribution of OB-type and Wolf-Rayet stars in proper
    motion relative to R136. VFTS682 is not an outlier, but
    its relative proper motion matches the expected value if it were indeed
    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to physical units
    assuming a distance of 50\,kpc. Right panel:  distribution of
    angles between the relative proper motion direction and the radial
    direction to the star. The error bars for VFTS682 are large, but
    the best value is in agreement with the hypothesis of dynamical
    ejection. In both
    panels, the dark blue histograms contain 317 
    stars with error smaller than $0.1\,\mathrm{mas \
      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
    lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
      yr^{-1}}$. \SdM{ Todo: [1] insert correct errorbars in right pabel, [2] make (a) linewidth, (b) line width of the hashed shading and (c)  figure height consistent. (You could try labeling the top x axis in the right panel in degrees.)  [3] fix normalization to be consistent in both pabels, [4] unit in axis label in righthand plot is not correct. Degrees but radian. [4] Legen: the rel error limits that we adopted are missing a unit. Milia arrsecond?    }}
  \label{fig:mu_dist}
\end{figure*}

@@

VFTS682 is identified with the source id 4657685637907503744 in the
\emph{Gaia} DR2 catalog  as a 15.65 mag star in the G band
\citep{gaia:16,brown:18}.   The number of visibility periods,
i.e. groups of observations separated from other groups by a gap of at
least four days, used in the astrometric solution is seventeen for this
star. The reported astrometric excess noise is zero.  These values
suggest that the \emph{Gaia} DR2  data for VFTS682 are
reliable. Gaia errors are correlated \citep[][]{lindengren:18}, but for this star the correlation coefficient is very small, $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$  = 0.023, where zero is for completely un-correlated data and one is for fully correlated data. We can thus safely treat the proper motion components as uncorrelated.

To determine the relative proper motion with respect to R136, we follow  \citet{lennon:18} to define the motion of the local frame of reference using the average proper motion of nearby stars with reliable astrometric data.  They select  bright ($G<17$) stars within 0.05 degrees of R136 with and exclude sources with proper motion errors greater than $0.01\masyr$ in both coordinates (see their  Sect.~2.1 for further discussion).

@@

The 30 doradus region was target of an a two-epoch photometric campaign with \emph{HST} providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}). Visually, the star is clearly isolated and does not appear to have any close neighbor. This by itself gives further confidence in ability of \emph{Gaia} to measure the proper motion reliably without being affected by crowding.

 \citet{platais:18} recently analyzed the HST data for candidate runaway stars. The brightest stars (V$<$14) are missing from the catalogue because of saturation effects, but the high extinction towards VFTS682 makes it faint enough to be included.
 Our star did not make the more conservative selection criteria for runaway stars adopted in that study. However, the measurements are useful here, since they provides a measurement of the proper motion of VFTS 682 that is completely independent from the \emph{Gaia} measurement.

@@

\begin{figure}%[tbp]
  \centering
  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
  \caption{\SdM{I don't see a reason why we would not show the HST data, except for wanting to hide it maybe.  Let's sit together and find a good way to plot both.  (Hope you still have the script to show the HST direction. )}  The red solid arrow indicates the proper motion of VFTS682. 
    relative to the region from \emph{Gaia} DR2, starting from the present day position of
    the star. The yellow arrows indicate the possible
    directions of projected motion within the \emph{Gaia} DR2 errors, and are extended
    backwards (dashed) to illustrate the uncertainty on the origin of the
    star. The length of the prolongations is proportional to the relative proper motion
    times the age of VFTS682 \citep[$1.0\pm0.2$\,Myr,][]{schneider:18}.
    % \todo{load color picture on background? Fig with HST data
    % available, but confusing...}
  }

>>>>

       \hline
           \multicolumn{5}{l}{\emph{Gaia's proper motion for VFTS 682
      and average of the region}} \\
      \hline
          &$\mu_\mathrm{RA}$&[$\masyr$] & $1.843\pm 0.070$
                     & (1) \\        
               &$\mu_\mathrm{DEC}$&[$\masyr$] & $0.786\pm 0.080$
                     &  (1) \\        
                 & $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$ &[$\masyr$] & $0.026$
                        & (1)  \\         
     %  \multicolumn{5}{l}{\emph{Gaia's average proper motion in the region}} \\         
       &$\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$&[$\masyr$] & $1.74\pm0.01$
                        & (2) \\
      &$\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$&[$\masyr$]
                & $0.70\pm0.02$ &  (2)\\
\hline

@@

\begin{figure*}[htbp]
  \centering
  \includegraphics[width=0.35\textwidth]{figures/dist_mu_region.pdf}
  \includegraphics[width=0.35\textwidth]{figures/angle}
  \caption{Left panel: distribution of OB-type and Wolf-Rayet stars in proper
    motion relative to R136. VFTS682 is not an outlier, but
    its relative proper motion matches the expected value if it were indeed
    ejected from R136 assuming an age of $1.0\pm0.2$\,Myr. The top axis shows the conversion to physical units
    assuming a distance of 50\,kpc. Right panel:  distribution of
    angles between the relative proper motion direction and the radial
    direction to the star. The error bars for VFTS682 are large, but
    the best value is in agreement with the hypothesis of dynamical
    ejection. In both
    panels, the dark blue histograms contain 317 
    stars with error smaller than $0.1\,\mathrm{mas \
      yr^{-1}}\simeq25\,\mathrm{km\ s^{-1}}$ at 50\,kpc, the
    lighter blue histograms contain 36 stars with errors smaller than $0.05\,\mathrm{mas \
      yr^{-1}}$. \SdM{ Todo: [1] insert correct errorbars in right pabel, [2] make (a) linewidth, (b) line width of the hashed shading and (c)  figure height consistent. (You could try labeling the top x axis in the right panel in degrees.)  [3] fix normalization to be consistent in both pabels, [4] unit in axis label in righthand plot is not correct. Degrees but radian. [4] Legen: the rel error limits that we adopted are missing a unit. Milia arrsecond?    }}
  \label{fig:mu_dist}
\end{figure*}

@@

VFTS682 is identified with the source id 4657685637907503744 in the
\emph{Gaia} DR2 catalog  as a 15.65 mag star in the G band
\citep{gaia:16,brown:18}.   The number of visibility periods,
i.e. groups of observations separated from other groups by a gap of at
least four days, used in the astrometric solution is seventeen for this
star. The reported astrometric excess noise is zero.  These values
suggest that the \emph{Gaia} DR2  data for VFTS682 are
reliable. Gaia errors are correlated \citep[][]{lindengren:18}, but for this star the correlation coefficient is very small, $\rho\,(\mu_\mathrm{RA}, \mu_\mathrm{DEC})$  = 0.026, where zero is for completely un-correlated data and one is for fully correlated data. We can thus safely treat the proper motion components as uncorrelated.

To determine the relative proper motion with respect to R136, we follow  \citet{lennon:18} to define the motion of the local frame of reference using the average proper motion of nearby stars with reliable astrometric data.  They select  bright ($G<17$) stars within 0.05 degrees of R136 and exclude sources with proper motion errors greater than $0.01\masyr$ in both coordinates (see their  Sect.~2.1 for further discussion).

@@

The 30 doradus region was target of an a two-epoch photometric campaign with HST providing observations in the F775W filter in October 2011 and October 2014 (GO-12499; P.I.: D.~J.~Lennon, \citealt{sabbi:13}). Visually, the star is clearly isolated and does not appear to have any close neighbor. This by itself gives further confidence in ability of \emph{Gaia} to measure the proper motion reliably without being affected by crowding.

 \citet{platais:18} recently analyzed the HST data for candidate
 OB-type runaway stars. The brightest stars (V$<$14) are missing from
 the catalogue because of saturation effects, but the high extinction
 towards VFTS682 makes it faint enough to be included.
 This star was not included in their study because of its WNh5
 spectral type and because it would not pass the more conservative
 selection criteria they adopted.
 However, the measurements are useful here, since they provides an
 estimate of the proper motion of VFTS 682 that is completely independent from the \emph{Gaia} measurement.

@@

\begin{figure}%[tbp]
  \centering
  \includegraphics[width=0.48\textwidth]{./figures/main_plot_good_gaia_only}  
  \caption{\SdM{I don't see a reason why we would not show the HST data, except for wanting to hide it maybe.  Let's sit together and find a good way to plot both.  (Hope you still have the script to show the HST direction. )}  The red solid arrow indicates the proper motion of VFTS682 
    relative to the region from \emph{Gaia} DR2, starting from the present day position of
    the star. The yellow arrows indicate the possible
    directions of projected motion within the \emph{Gaia} DR2 errors, and are extended
    backwards (dashed) to illustrate the uncertainty on the origin of the
    star. The length of the prolongations is proportional to the relative proper motion
    times the age of VFTS682 \citep[$1.0\pm0.2$\,Myr,][]{schneider:18}.
    % \todo{load color picture on background? Fig with HST data
    % available, but confusing...}
  }
